---
title: "Data Wrangling"
author: "Miguel Peñate"
#date: "`r Sys.Date()`"
format:
  html:
    code-fold: true
    self-contained: false
---
>Data wrangling (Cleaning & Preparation) is the important process of preparing data for analysis and the foundation of exploratory data analysis.

In this section we will work with the data about: [Patents](https://cloud.tuhh.de/index.php/s/sGZKmdfLfdZBFK9)

### QUESTION 1

"The table shows the top 10 US companies with the most assigned/granted patents."

```{r plot, fig.width=10, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

setwd("C:/Users/migue/Desktop/Hamburg/TUHH/ss24-bdsb-miguelp98/01_raw_data")
# Cargar librerías necesarias
library(tidyverse)
library(vroom)
library(data.table)
library(tictoc)
library(kableExtra)

# Definir los tipos de columnas para cada archivo

# assignee.tsv
col_types_assignee <- list(
  id = col_character(),
  type = col_character(),
  organization = col_character()
)

# patent.tsv
col_types_patent <- list(
  id = col_character(),
  date = col_date("%Y-%m-%d"),
  num_claims = col_integer()
)

# patent_assignee.tsv
col_types_patent_assignee <- list(
  patent_id = col_character(),
  assignee_id = col_character()
)

# uspc.tsv
col_types_uspc <- list(
  patent_id = col_character(),
  mainclass_id = col_character(),
  sequence = col_integer()
)

# Importar los datos

# assignee.tsv
assignee_tbl <- vroom(
  file       = "assignee.tsv", 
  delim      = "\t", 
  col_types  = col_types_assignee,
  na         = c("", "NA", "NULL")
)
setDT(assignee_tbl)

# patent.tsv
patent_tbl <- vroom(
  file       = "patent.tsv", 
  delim      = "\t", 
  col_types  = col_types_patent,
  na         = c("", "NA", "NULL")
)
setDT(patent_tbl)

# patent_assignee.tsv
patent_assignee_tbl <- vroom(
  file       = "patent_assignee.tsv", 
  delim      = "\t", 
  col_types  = col_types_patent_assignee,
  na         = c("", "NA", "NULL")
)
setDT(patent_assignee_tbl)

# uspc.tsv
uspc_tbl <- vroom(
  file       = "uspc.tsv", 
  delim      = "\t", 
  col_types  = col_types_uspc,
  na         = c("", "NA", "NULL")
)
setDT(uspc_tbl)

# Renombrar columnas para consistencia
setnames(patent_assignee_tbl, old = c("patent_id", "assignee_id"), new = c("id", "assignee_id"))
setnames(uspc_tbl, old = "patent_id", new = "id")

# Combinar los datos paso a paso
combined_data <- merge(patent_tbl, patent_assignee_tbl, by = "id")
combined_data <- merge(combined_data, assignee_tbl, by.x = "assignee_id", by.y = "id")
combined_data <- merge(combined_data, uspc_tbl, by = "id")

# Filtrar las empresas americanas
us_assignees <- combined_data[type %in% c(2, 4, 6, 8, 9)]

# Visualizar una muestra de los datos combinados
#us_assignees %>% glimpse()

# Contar las patentes por organización
patent_count <- us_assignees[, .N, by = organization][order(-N)][1:10]

# Mostrar el resultado en formato de tabla
kable(patent_count, format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```


---

### QUESTION 2

"The table lists the top 10 US companies with the most new granted patents for August 2014"

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Filtrar patentes de agosto de 2014 y contar por organización
august_patents <- us_assignees[format(date, "%Y-%m") == "2014-08", .N, by = organization][order(-N)][1:10]

# Mostrar el resultado en formato de tabla
kable(august_patents, format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

---

### QUESTION 3

"The table shows the top 5 USPTO tech main classes for the top 10 companies worldwide with the most patents."

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Filtrar datos y contar las principales clases tecnológicas
top_tech_sectors <- combined_data[, .N, by = mainclass_id][order(-N)][1:5]

# Mostrar el resultado en formato de tabla
kable(top_tech_sectors, format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```